---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# permAprox

permAprox ("**perm**utation p-value **ap**p**rox**imation") is an R package to compute empirical and approximated p-values from permutation tests, especially useful when the number of permutations is small and zero p-values should be strictly avoided. It offers approximation using the Gamma distribution or the Generalized Pareto Distribution (GPD) fitted to the tail of the permutation distribution.

## Early version disclaimer

**Note:** This is an early version of the permAprox package and is still under active development. While the core functionality is in place, the interface, parameter settings, and behavior of certain functions may change in future versions. Users are encouraged to use the package with care and avoid relying on its current structure for production workflows.

## Installation

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("stefpeschel/permAprox")
```

## Features

- Computes empirical p-values for permutation tests
- Approximates small p-values using the Gamma or GPD distributions
- Strictly avoids zero p-values
- Includes multiple testing correction support
- Control objects allow fine-tuning of each fitting method

## Usage

```{r}
library(permAprox)

# Simulated example
set.seed(12345)
obs <- c(2.0, 3.0, 4.0, 5.0)
perm <- matrix(rnorm(4000), nrow = 4)

# Empirical p-values
res_emp <- compute_p_values(obs_stats = obs,
                            perm_stats = perm,
                            method = "empirical")

# Gamma approximation
gamma_ctrl <- make_gamma_ctrl(gof_test = "none")
res_gamma <- compute_p_values(obs_stats = obs,
                              perm_stats = perm,
                              method = "gamma",
                              gamma_ctrl = gamma_ctrl)

# GPD approximation
res_gpd <- compute_p_values(obs_stats = obs,
                            perm_stats = perm,
                            method = "gpd")

# GPD with constraint
gpd_ctrl <- make_gpd_ctrl(constraint = "support_at_obs")
res_gpd_constr <- compute_p_values(obs_stats = obs,
                                   perm_stats = perm,
                                   method = "gpd",
                                   gpd_ctrl = gpd_ctrl)

# Compare unadjusted p-values
p_values <- data.frame(empirical = res_emp$p_unadjusted,
                       gamma = res_gamma$p_unadjusted,
                       gpd = res_gpd$p_unadjusted,
                       gpd_constr = res_gpd_constr$p_unadjusted)

print(p_values)
```


## Documentation

- The main function is `compute_p_values()`
- Control objects can be created using:
  - `make_gpd_ctrl()` for GPD fitting
  - `make_gamma_ctrl()` for Gamma fitting
  - `make_adjust_ctrl()` for multiple testing correction

Use `?compute_p_values` in R to see the full documentation and available options.

## License

This package is licensed under GPL-3.
